### 2024-12-10 Draft_Cleaning.Rmd
split by rivers for PCA, clean PCA plots, got rid of unneeded doc

---
title: "DRAFT_Cleaning"
output: html_document
date: "2024-12-04"
---

```{r setup, include=FALSE}
library(readr)
library(tidyverse)
library(lubridate)
library(caret)
library(ggfortify)
library(ggformula)
library(DiagrammeR)
```

Read in the data
```{r}
dredge_data<-read_csv("DredgeData.csv", show_col_types = FALSE)
survey_data<-read_csv("SurveyData.csv", show_col_types = FALSE)
gage_data<-read_csv("UMR_IWW_1999_2021.csv", show_col_types = FALSE)

### The gage data has multiple records across different gages per day, we need to
### combine the rows by date so that there's only one record for date across all gages.
### Since we're going to be doing a lot of work with the date column, lets format it as 
### a date object and arrange by DATE_START and remove the old Date column
gage_data<- gage_data |>
  group_by(Date)|>
   summarize(across(everything(),~first(na.omit(.)), .names = "{.col}"))|>
  mutate(DATE_START= as_date(dmy(Date)))|>
    arrange(DATE_START)|>
  select(-Date)

### Some of the gage data has observation gaps where the sensors may not have been
### working, lets interpolate gage values where there's data before and after a gap
### with a maximum window of 2
gage_data<-gage_data|>
    mutate(across(c(-DATE_START),
                ~ ifelse(is.na(.),
                         coalesce((lag(.) + lead(.)) / 2, lag(.), lead(.)),
                         .)))
```

Now lets clean the survey data. I only want conditional survey records between
1999 to 2000, these are going to act as my background data. I want to combine
these records with my dredge data. I'm also going to filter by river, remove
pools that are not of interest, clean up the SURVEYJOBIDPK, and create some 
columns to support my join. Dates are also formatted as date objects and some
columns are renamed to make it easier to join my surveys to my dredge data.
Finally, select the columns we need for our analysis to join to our dredging 
data. 
```{r}
survey_data<- survey_data |>
  filter(SURVEYTYPE == "conditional",
         SURVEY_YEAR >= 1999 & SURVEY_YEAR <= 2022,
          RIVER == "Mississippi_River" | RIVER == "Illinois_Waterway",
         !POOL %in% c("LP","CS"))|>
  mutate(VOLUMEDREDGED = 0,
         DREDGINGPURPOSE = "None",
         DREDGE_TYPE = "None",
         LABOR_TYPE = LABOR_SOURCE,
         EXECYEAR = SURVEY_YEAR,
         FEATURENAME = SDSFEATURENAME,
         DATE_START = as.character(DATE_START),
         DATE_START = as_date(DATE_START)
         )|>
  select(EXECYEAR,DATE_START,DREDGINGPURPOSE,VOLUMEDREDGED, RIVER,POOL, DREDGE_TYPE,
         UP_RIV_MIL,DN_RIV_MIL)
```

Lets clean the dredge data, we only want records that are channel dredging events
and since the dredge records only go back to 1999, we just need to filter out data beyond 2022.
```{r}
dredge_data <- dredge_data |>
   filter(
          DREDGINGPURPOSE != "harbor",
           EXECYEAR <= 2022)|>
  mutate(
     DATE_START = as_date(mdy_hm(DATE_START)))|>
  select(EXECYEAR,DREDGINGPURPOSE, VOLUMEDREDGED, RIVER,POOL, DREDGE_TYPE,
         DATE_START, UP_RIV_MIL,DN_RIV_MIL)
```

Combine the dredge data and survey data into 1 table. The survey data will act as background/non-dredge data.
```{r}
combined_data <- bind_rows(survey_data, dredge_data)

combined_data <-combined_data |>
     filter(DATE_START >= as.Date("1999-04-01") &
         DATE_START <= as.Date("2021-09-30"))

combined_data <-combined_data |>
  mutate(lead_7 = (DATE_START - 7),
         lead_14 = (DATE_START - 14))
Nas<- combined_data|>
  filter(if_any(everything(),is.na))

gage_survey <- combined_data |>
  left_join(gage_data, by = "DATE_START")

na<-gage_survey |>
  select(everything()) |>
  summarise_all(list(~sum(is.na(.))))
na

NA_by_row<- gage_survey |>
  is.na() |>
  rowSums()

gf_histogram(~NA_by_row)

gage_survey_filt <- gage_survey |>
  filter(complete.cases(gage_survey))

gage_survey_full <- gage_survey_filt|>  
  left_join(gage_data, by = c("lead_7" = "DATE_START"), suffix = c("", "_7"))|>
  left_join(gage_data, by = c("lead_14" = "DATE_START"), suffix = c("","_14"))

Nas<- gage_survey_full|>
  select(-lead_7,-lead_14)|>
  filter(if_any(everything(),is.na))

gage_survey_cleaned <- gage_survey_full |>
  filter(complete.cases(gage_survey_full))

final_na<-gage_survey_cleaned|>
 select(-lead_7,-lead_14)|>
  filter(if_any(everything(),is.na))

```

Create factored season variable that could be used in analysis
```{r}
date_to_season <- function(date) {
  month <- as.numeric(format(date, "%m"))
  day <- as.numeric(format(date, "%d"))
  # Define the seasons based on date ranges
  if ((month == 3 && day >= 20) || (month >= 4 && month <= 6) || (month == 6 && day <= 20)) {
    return("Spring")
  } else if ((month == 6 && day >= 21) || (month >= 7 && month <= 9) || (month == 9 && day <= 21)) {
    return("Summer")
  } else if ((month == 9 && day >= 22) || (month >= 10 && month <= 12) || (month == 12 && day <= 20)) {
    return("Fall")
  } else {
    return("Winter")
  }
}

gage_survey_cleaned <- gage_survey_cleaned |>
  mutate("Season" = sapply(DATE_START,date_to_season))

```

```{r}
gage_survey_PCA <- gage_survey_cleaned |>
select(where(is.numeric), -EXECYEAR)

dredge_numeric<-gage_survey_cleaned|>
   select(where(is.numeric), -EXECYEAR)
correlations <-cor(dredge_numeric, use = "pairwise.complete.obs")
corrplot::corrplot(correlations)

dredge_prcomp <- gage_survey_PCA|>
  select(-VOLUMEDREDGED)

dredge_PCA = prcomp(dredge_prcomp, scale = T)

biplot(dredge_PCA)

dredge <-gage_survey_cleaned |>
  filter(DREDGINGPURPOSE != "None")
  hist(dredge$VOLUMEDREDGED)

  median(dredge$VOLUMEDREDGED)

hist(gage_survey_cleaned$VOLUMEDREDGED)

plot(dredge_PCA,  type="l")

autoplot(dredge_PCA,data = gage_survey_cleaned, 
         color = 'SEASON',loadings = TRUE,
         loadings.colour = 'black',
         loadings.label = TRUE, loadings.label.size = 4, 
         loadings.label.colour = "blue")

variable_contributions<- dredge_PCA$rotation
head(variable_contributions)

set.seed(12)
ctrl = trainControl(method = "cv", number = 5)

fit_pca = train(VOLUMEDREDGED ~ .,
            data = gage_survey_PCA,
            method = "glm",
            preProc = c("pca"),
            trControl = ctrl)
fit_pca

summary(fit_pca$finalModel)
varImp(fit_pca)
variable_contributions = fit_pca$preProcess$rotation
head(variable_contributions)

# data.frame(variable_contributions) |>
#   mutate(RIVER = gage_survey_cleaned$RIVER)
#   gf_point(PC2 ~ PC1)

data.frame(variable_contributions) |>
  gf_text(PC2 ~ PC1,
          label = row.names(variable_contributions))

data.frame(variable_contributions) |>
  filter(abs(PC1) > 0.11)

predictor_matrix <- gage_survey_PCA |>
  dplyr::select(-VOLUMEDREDGED) |> # Remove the response variable
  as.matrix() |>
  scale() #centers and scales

# The symbol %*% does matrix multiplication
data_in_PC_space = predictor_matrix %*% variable_contributions

data_in_PC_space = data.frame(data_in_PC_space,
                              River = gage_survey_cleaned$RIVER,
                              Pool = gage_survey_cleaned$POOL,
                              DREDGE_TYPE = gage_survey_cleaned$DREDGINGPURPOSE,
                              Season = gage_survey_cleaned$Season)

data.frame(data_in_PC_space)|>
  gf_point(PC2 ~ PC1, col = ~ Season)

data.frame(data_in_PC_space)|>
  gf_point(PC2 ~ PC1, col = ~DREDGE_TYPE)

### Scree Plot
data_in_PC_space = predictor_matrix %*% variable_contributions
var_explained = apply(data_in_PC_space,2,var)
pve = var_explained/sum(var_explained)
pve

cumsum(pve)

var_df <- data.frame(pve, PC = 1:length(pve),
                     cum_var = cumsum(pve))

var_df|>
  gf_point(cum_var ~ PC) |>
  gf_line(cum_var ~ PC)

pca_scores <-dredge_PCA$x
pca_loadings <- dredge_PCA$rotation

```

###xgBoost
```{r}
library(xgboost)
gage_xgb <- gage_survey_cleaned |>
  select(-lead_7, -lead_14,-DATE_START,
         -SURVEYJOBIDPK,-EXECYEAR,-DREDGE_TYPE,
         -DREDGINGPURPOSE)|>
    mutate_if(is.character, factor)

ctrl = trainControl(method = "cv", number = 5) 
fit_dredge = train(VOLUMEDREDGED ~ .,
                 data = gage_xgb,
                 method = "xgbTree",
               tuneGrid = expand.grid(nrounds = c (15,25,50,100), 
                                              max_depth = 1:10,
                                        eta =c(0.3), gamma = 0,
                                        colsample_bytree = 1,
                                        min_child_weight = 1, subsample = 1),
                 verbosity = 0,
                 trControl = ctrl,
                 na.action = na.exclude)

varImp(fit_dredge)
xgb.plot.tree(model = fit_dredge$finalModel, trees = 0)

fit_dredge

import_matrix = xgb.importance(model = fit_dredge$finalModel)
import_matrix
xgb.plot.importance(import_matrix)
```
###Cross Validation of Model
```{r}
set.seed(123)
n = nrow(gage_xgb)
n_outer_folds = 5

cv_pred = vector(length = n)

groups = rep(1:n_outer_folds, length = n)
cv_groups = sample(groups, n)

for(ii in 1:n_outer_folds){
  in_test = (cv_groups == ii)
  in_train = (cv_groups != ii)

  ctrl = trainControl(method = "cv", number = 5)
  fit_dredge_DCV = train(VOLUMEDREDGED ~ .,
                       data = gage_xgb[in_train, ],
                       method = "xgbTree",
                       tuneGrid = expand.grid(nrounds = c (15,25,50,100), 
                                              max_depth = 1:10,
                                        eta =c(0.3), gamma = 0,
                                        colsample_bytree = 1,
                                        min_child_weight = 1, subsample = 1),
                       verbosity = 0,
                       trControl = ctrl,
                       na.action = na.exclude)

  # "Frankenstein" vector of predictions
  cv_pred[in_test] = predict(fit_dredge_DCV,
                             newdata = gage_xgb[in_test, ])

}

##Og
train<-gage_xgb[in_train, ]
best_tune<-fit_dredge_DCV$results
best_tune
confusion_matrix <- confusionMatrix(cv_pred[in_test], gage_xgb[in_test, ])
print(confusion_matrix)

# I'll use RMSE to assess the model, because that's the metric I used to select the model.
# The model with the best RMSE may or may not be the one with the best MAE.
# RMSE
sqrt(mean((cv_pred - gage_xgb$VOLUMEDREDGED)^2)) # This is the most honest
                                         # assessment of accuracy
```

## Well that didnt work! Lets try modeling between the two river systems

### Cleaning the data 
```{r}
combined_data <- bind_rows(survey_Data, dredge_data_cleaning)

combined_data <-combined_data |>
     filter(DATE_START >= as.Date("1999-04-01") &
         DATE_START <= as.Date("2021-09-30"))

combined_data <-combined_data |>
  mutate(lead_7 = (DATE_START - 7),
         lead_14 = (DATE_START - 14))
Nas<- combined_data|>
  filter(if_any(everything(),is.na))

gage_survey <- combined_data |>
  left_join(gage_data, by = "DATE_START")

Illinois_River <- gage_survey |>
  select(SURVEYJOBIDPK,EXECYEAR,DATE_START,DREDGINGPURPOSE,VOLUMEDREDGED,RIVER,
         POOL, DREDGE_TYPE,UP_RIV_MIL, DN_RIV_MIL,lead_7,lead_14,BEAI2,HAVI2,
         HNYI2,IL03, IL03_Tail, IL04, IL04_Tail, IL05, IL05_Tail,IL06, IL06_Tail,
         IL07, IL07_Tail, IL08, IL08_Tail,KNGI2, MORI2, PIAI2) |>
  filter(RIVER == "Illinois_Waterway")

Miss_River <- gage_survey |>
  select(SURVEYJOBIDPK,EXECYEAR,DATE_START,DREDGINGPURPOSE,VOLUMEDREDGED,RIVER,
         POOL, DREDGE_TYPE,UP_RIV_MIL, DN_RIV_MIL,lead_7,lead_14,BRLI4,CMMI4, DBQI4,
         FAII4, KHBI2,MI11,MI11_Tail,MI12,MI12_Tail,MI13,MI13_Tail,MI14,MI14_Tail,
         MI15,MI15_Tail,MI16,MI16_Tail,MI17,MI17_Tail,MI18,MI18_Tail,MI19,MI19_Tail,
         MI20,MI20_Tail,MI21,MI21_Tail,MI22,MI22_Tail, MUSI4,UINI2) |>
  filter(RIVER == "Mississippi_River")

Miss_full <- Miss_River|>  
  left_join(gage_data, by = c("lead_7" = "DATE_START"), suffix = c("", "_7"))|>
  left_join(gage_data, by = c("lead_14" = "DATE_START"), suffix = c("","_14"))

IWW_full <- Illinois_River|>  
  left_join(gage_data, by = c("lead_7" = "DATE_START"), suffix = c("", "_7"))|>
  left_join(gage_data, by = c("lead_14" = "DATE_START"), suffix = c("","_14"))

Miss_Cleaned <- Miss_full |>
  filter(complete.cases(Miss_full))

IWW_Cleaned <- IWW_full |>
  filter(complete.cases(IWW_full))

IWW_Cleaned <- IWW_Cleaned |>
  mutate("Season" = sapply(DATE_START,date_to_season))

Miss_Cleaned <- Miss_Cleaned |>
  mutate("Season" = sapply(DATE_START,date_to_season))

```

## PCA
```{r}

```

## xgboost split by river
#Miss
```{r}
Miss_xgb <-Miss_Cleaned |>
  select(-lead_7, -lead_14,-DATE_START,
         -SURVEYJOBIDPK,-EXECYEAR,-DREDGE_TYPE,
         -DREDGINGPURPOSE, -RIVER)|>
    mutate_if(is.character, factor)

ctrl = trainControl(method = "cv", number = 10) 
Miss_xgb_fit = train(VOLUMEDREDGED ~ .,
                 data = Miss_xgb,
                 method = "xgbTree",
               tuneGrid = expand.grid(nrounds = c(15,25,50,100), 
                                              max_depth = 1:10,
                                        eta =c(0.3), gamma = 0,
                                        colsample_bytree = 1,
                                        min_child_weight = 1, subsample = 1),
                 verbosity = 0,
                 trControl = ctrl,
                 na.action = na.exclude)

varImp(Miss_xgb_fit)
xgb.plot.tree(model = Miss_xgb_fit$finalModel, trees = 0)

Miss_xgb_fit

import_matrix = xgb.importance(model = Miss_xgb_fit$finalModel)
import_matrix
xgb.plot.importance(import_matrix)
```

#IWW
```{r}
IWW_xgb <-IWW_Cleaned |>
  select(-lead_7, -lead_14,-DATE_START,
         -SURVEYJOBIDPK,-EXECYEAR,-DREDGE_TYPE,
         -DREDGINGPURPOSE, -RIVER)|>
    mutate_if(is.character, factor)

ctrl = trainControl(method = "cv", number = 10) 
IWW_xgb_fit = train(VOLUMEDREDGED ~ .,
                 data = IWW_xgb,
                 method = "xgbTree",
               tuneGrid = expand.grid(nrounds = c(15,25,50,100), 
                                              max_depth = 1:10,
                                        eta =c(0.3), gamma = 0,
                                        colsample_bytree = 1,
                                        min_child_weight = 1, subsample = 1),
                 verbosity = 0,
                 trControl = ctrl,
                 na.action = na.exclude)

varImp(IWW_xgb_fit)
xgb.plot.tree(model = IWW_xgb_fit$finalModel, trees = 0)

IWW_xgb_fit

import_matrix = xgb.importance(model = IWW_xgb_fit$finalModel)
import_matrix
xgb.plot.importance(import_matrix)
```