### 2025-06-26 Draft_Cleaning.Rmd
testing different LSTM configs

---
title: "DRAFT_Cleaning"
output: html_document
date: "2024-12-04"
---

```{r setup, include=FALSE}
library(readr)
library(tidyverse)
library(lubridate)
library(caret)
library(ggfortify)
library(ggformula)
library(DiagrammeR)
library(gridExtra)
library(keras)
library(tensorflow)

### Key Functions 
# Function to plot PCA
plot_pca <- function(PCAvalues, top_loadings, PC1_explained, PC2_explained, grouping_var) {

  p <- ggplot(PCAvalues, aes(x = "PC1", y = "PC2", colour = grouping_var)) +
    geom_point(size = 2) +  # Scatter plot
    geom_segment(data = top_loadings, aes(x = 0, y = 0, xend = PC1 * 75, yend = PC2 * 75),
                 arrow = arrow(length = unit(1/2, "picas")), color = "black") +  # Vectors for top loadings
    annotate("text", x = top_loadings$PC1 * 80, y = top_loadings$PC2 * 80,
             label = top_loadings$Variables, hjust = 0.2, vjust = 0.4) +  # Labels for top loadings
    theme_minimal() +
    labs(title = paste("PCA Plot with Top 10 Loadings for PC1 and PC2 - Grouped by", grouping_var),
         x = paste("PC1 (", round(PC1_explained, 1), "%)", sep = ""),
         y = paste("PC2 (", round(PC2_explained, 1), "%)", sep = ""))

  return(p)
}

#Function to calculate season from date 
date_to_season <- function(date) {
  month <- as.numeric(format(date, "%m"))
  day <- as.numeric(format(date, "%d"))
  # Define the seasons based on date ranges
  if ((month == 3 && day >= 20) || (month >= 4 && month <= 6) || (month == 6 && day <= 20)) {
    return("Spring")
  } else if ((month == 6 && day >= 21) || (month >= 7 && month <= 9) || (month == 9 && day <= 21)) {
    return("Summer")
  } else if ((month == 9 && day >= 22) || (month >= 10 && month <= 12) || (month == 12 && day <= 20)) {
    return("Fall")
  } else {
    return("Winter")
  }
}

```

### Read in the data
```{r}
CSAT_dredge = read_csv("C:/workspace/DRAFT/CSAT_Data/Dredged_Date_Vol.csv",show_col_types = FALSE)
dredge_data<-read_csv("DredgeData.csv", show_col_types = FALSE)
survey_data<-read_csv("SurveyData.csv", show_col_types = FALSE)
gage_data<-read_csv("UMR_IWW_1999_2024.csv", show_col_types = FALSE)
```

### Clean the gage data 
The gage data has multiple records across different gages per day, we need to
combine the rows by date so that there's only one record for date across all gages.
Since we're going to be doing a lot of work with the date column, lets format it as 
a date object and arrange by DATE_START and remove the old Date column
```{r}
gage_data<- gage_data |>
  group_by(Date)|>
   summarize(across(everything(),~first(na.omit(.)), .names = "{.col}"))|>
  mutate(DATE_START= as_date(dmy(Date)))|>
    arrange(DATE_START)|>
  select(-Date)

### Some of the gage data has observation gaps where the sensors may not have been
### working, lets interpolate gage values where there's data before and after a gap
### with a maximum window of 2
gage_data<-gage_data|>
    mutate(across(c(-DATE_START),
                ~ ifelse(is.na(.),
                         coalesce((lag(.) + lead(.)) / 2, lag(.), lead(.)),
                         .)))
```

### Cleaning the data
Now lets clean the survey data. I only want conditional survey records between
1999 to 2022, these are going to act as my background data. I want to combine
these records with my dredge data. I'm also going to filter by river, remove
pools that are not of interest, and create some 
columns to support my join. Dates are also formatted as date objects and some
columns are renamed to make it easier to join my surveys to my dredge data.
Finally, select the columns we need for our analysis to join to our dredging 
data. 
```{r}
survey_data<- survey_data |>
  filter(SURVEYTYPE == "conditional",
         SURVEY_YEAR >= 1999 & SURVEY_YEAR <= 2024,
          RIVER == "Mississippi_River" | RIVER == "Illinois_Waterway",
         !POOL %in% c("LP","CS"))|>
  mutate(VOLUMEDREDGED = 0,
         EXECYEAR = SURVEY_YEAR,
         DATE_START = as.character(DATE_START),
         DATE_START = as_date(DATE_START)
         )|>
  select(EXECYEAR,DATE_START,VOLUMEDREDGED, RIVER,POOL, 
         UP_RIV_MIL,DN_RIV_MIL)
```

```{r}
CSAT_dredge_clean <- CSAT_dredge |>
  mutate(VOLUMEDREDGED = VolumeChange,
         EXECYEAR = year(Date),
         DATE_START = SurveyDate,
         RIVER = case_when(River == "UM" ~"Mississippi_River",
                           River =="IL" ~ "Illinois_Waterway"), 
         POOL = Pool,
         UP_RIV_MIL = mileagestart,
         DN_RIV_MIL = mileageend)|>
  select(EXECYEAR,DATE_START,VOLUMEDREDGED,RIVER,POOL, 
         UP_RIV_MIL,DN_RIV_MIL)
```

Lets clean the dredge data, we only want records that are channel dredging events
and since the dredge records only go back to 1999, we just need to filter out data beyond 2022.
```{r}
dredge_data <- dredge_data |>
   filter(
          DREDGINGPURPOSE != "harbor",
           EXECYEAR <= 2024)|>
  mutate(
     DATE_START = as_date(mdy_hm(DATE_START)))|>
  select(EXECYEAR,DREDGINGPURPOSE, VOLUMEDREDGED, RIVER,POOL, DREDGE_TYPE,
         DATE_START, UP_RIV_MIL,DN_RIV_MIL)
```

Combine the dredge data and survey data into 1 table. The survey data will act as background/non-dredge data.
```{r}
combined_data <- bind_rows(survey_data, CSAT_dredge_clean)

combined_data <-combined_data |>
     filter(DATE_START >= as.Date("1999-04-01") &
         DATE_START <= as.Date("2024-12-31")) |>
           mutate( 
          #DREDGINGPURPOSE = case_when(DREDGINGPURPOSE == "None" ~ "No Dredging",
          #                             DREDGINGPURPOSE == "channel" ~ "Dredged"),
                  week_of_year = week(DATE_START))

combined_data <-combined_data |>
  mutate(lead_7 = (DATE_START - 7),
         lead_14 = (DATE_START - 14))
```

Check for NAs for PCA analysis. 
```{r}
Nas<- combined_data|>
  filter(if_any(everything(),is.na))

gage_survey <- combined_data |>
  left_join(gage_data, by = "DATE_START")

na<-gage_survey |>
  select(everything()) |>
  summarise_all(list(~sum(is.na(.))))
na

NA_by_row<- gage_survey |>
  is.na() |>
  rowSums()

gf_histogram(~NA_by_row)

gage_survey_filt <- gage_survey |>
  filter(complete.cases(gage_survey))

gage_survey_full = gage_survey_filt
write_csv(gage_survey_filt,"Gage_Survey_Dredge.csv")
```

Join the gage data by the 7 day prior observations and by the 14 day prior observations
```{r}
gage_survey_filt <- gage_survey_filt |>
  mutate(
    lead_7 = as.Date(lead_7),
    lead_14 = as.Date(lead_14)
  )

gage_data <- gage_data |>
  mutate(DATE_START = as.Date(DATE_START))

gage_survey_full <- gage_survey_filt|>  
  left_join(gage_data, by = c("lead_7" = "DATE_START"), suffix = c("", "_7"))|>
  left_join(gage_data, by = c("lead_14" = "DATE_START"), suffix = c("","_14"))
```

Check again for NAs, remove any remaining NA rows. 
```{r}
Nas<- gage_survey_full|>
  select(-lead_7,-lead_14)|>
  filter(if_any(everything(),is.na))

gage_survey_cleaned <- gage_survey_full |>
  filter(complete.cases(gage_survey_full))

final_na<-gage_survey_cleaned|>
 select(-lead_7,-lead_14)|>
  filter(if_any(everything(),is.na))

```

Create factored season variable that could be used in analysis
```{r}
gage_survey_cleaned <- gage_survey_cleaned |>
  mutate("Season" = sapply(DATE_START,date_to_season))

```

Lets make some plots of the data!
```{r}
gages_pivoted <- gage_data|>
    pivot_longer(cols = -DATE_START, 
               names_to = "gage", 
               values_to = "value")

gage_survey_cleaned |>
  ggplot(aes(y=VOLUMEDREDGED, x=DATE_START))+
  geom_line()

gages_pivoted |>
  ggplot(aes(y = value, x = DATE_START, color = gage, group = gage)) +
  geom_line(size = 1) +  # Adjust line thickness
  scale_color_viridis_d() +  # A color scale that's visually appealing
  labs(
    title = "Time Series of Gage Values",
    x = "Date",
    y = "Value",
    color = "Gage"
  ) +  # Adding titles and labels
  theme_minimal() +  # Use a minimal theme
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),  # Rotate x-axis labels for better readability
    axis.title = element_text(size = 12),  # Increase title font size
    legend.position = "right",  # Position the legend on the right
    legend.title = element_text(size = 12),
    legend.text = element_text(size = 10)
  )

gages_pivoted |>
  ggplot(aes())
```

```{r}
gage_survey |>
  filter(RIVER == "Mississippi_River")|>
  ggplot(aes(y=VOLUMEDREDGED, x=DATE_START,color = POOL))+
  geom_point()+
  labs(title = "Miss Dredging")

gage_survey |>
  filter(RIVER == "Illinois_Waterway")|>
  ggplot(aes(y=VOLUMEDREDGED, x=DATE_START, color = POOL))+
  geom_point()+
  labs(title = "IWW Dredging")

```

Now on to analysis!

To first explore the variables, we will use PCA.
### PCA
PCA using prcomp
```{r}
gage_survey_PCA <- gage_survey_cleaned |>
select(where(is.numeric), -EXECYEAR)

dredge_prcomp <- gage_survey_PCA|>
  select(-VOLUMEDREDGED)

dredge_PCA = prcomp(dredge_prcomp, scale = T)

plot(dredge_PCA,  type="l")

a<-autoplot(dredge_PCA,data = gage_survey_cleaned, 
         color = 'Season',loadings = FALSE)+
  ggtitle("Season")

b<-autoplot(dredge_PCA,data = gage_survey_cleaned, 
         color = 'POOL',loadings= FALSE)+
  ggtitle("River Pool")+
    guides(color = guide_legend(ncol = 2))

c<-autoplot(dredge_PCA,data = gage_survey_cleaned, 
         color = 'RIVER',loadings = FALSE)+
  ggtitle("River")

# d<-autoplot(dredge_PCA,data = gage_survey_cleaned, 
#          color = 'DREDGINGPURPOSE',loadings = FALSE)+
#         ggtitle("Dredging Purpose")+
#   labs(color = "Dredging?")

e<-autoplot(dredge_PCA,data = gage_survey_cleaned, 
         color = 'week_of_year',loadings = FALSE)+
        ggtitle("Week ")+
  labs(color = "Dredging?")

ggsave("Initial_Full_PCA.png",grid.arrange(a,b,c,e,ncol = 2),width = 10, height = 8)
variable_contributions<- dredge_PCA$rotation
head(variable_contributions)
```

PCA using Caret
```{r}
set.seed(12)
ctrl = trainControl(method = "cv", number = 5)

fit_pca = caret::train(VOLUMEDREDGED ~ .,
            data = gage_survey_PCA,
            method = "glm",
            preProc = c("pca"),
            trControl = ctrl)
fit_pca

summary(fit_pca$finalModel)
varImp(fit_pca)
variable_contributions = fit_pca$preProcess$rotation
head(variable_contributions)

data.frame(variable_contributions) |>
  gf_text(PC2 ~ PC1,
          label = row.names(variable_contributions))

data.frame(variable_contributions) |>
  filter(abs(PC1) > 0.11)

predictor_matrix <- gage_survey_PCA |>
  dplyr::select(-VOLUMEDREDGED) |> # Remove the response variable
  as.matrix() |>
  scale() #centers and scales

# The symbol %*% does matrix multiplication
data_in_PC_space = predictor_matrix %*% variable_contributions

data_in_PC_space = data.frame(data_in_PC_space,
                              River = gage_survey_cleaned$RIVER,
                              Pool = gage_survey_cleaned$POOL,
                              Season = gage_survey_cleaned$Season,
                              Dredge_Volume = gage_survey_cleaned$VOLUMEDREDGED)

data.frame(data_in_PC_space)|>
  gf_point(PC2 ~ PC1, col = ~ Season)

# data.frame(data_in_PC_space)|>
#   gf_point(PC2 ~ PC1, col = ~DREDGE_TYPE)

### Scree Plot
data_in_PC_space = predictor_matrix %*% variable_contributions
var_explained = apply(data_in_PC_space,2,var)
pve = var_explained/sum(var_explained)
pve

cumsum(pve)

var_df <- data.frame(pve, PC = 1:length(pve),
                     cum_var = cumsum(pve))

var_df|>
  gf_point(cum_var ~ PC) |>
  gf_line(cum_var ~ PC)
```

### xgboost og
```{r}
library(xgboost)
gage_survey_sort <- gage_survey_cleaned|>
  arrange(DATE_START)  # If DATE_START was dropped, sort gage_survey_cleaned first

gage_xgb <- gage_survey_sort |>
  select(-lead_7, -lead_14,
    -DATE_START,
         -EXECYEAR)|>
    mutate_if(is.character, factor)

gage_xgb_scaled <- gage_xgb |>
  mutate(across(where(is.numeric), ~ scale(.) |> 
                  as.data.frame() |>
                  pull(1), .names = "{.col}"))

ctrl <- trainControl(
  method = "timeslice",
  initialWindow = floor(0.7 * nrow(gage_xgb_scaled)),  # 70% of data for initial training
  horizon = 1,               # Predict 1 step ahead
  fixedWindow = TRUE,        # Rolling window
  verboseIter = TRUE
)

fit_dredge = caret::train(VOLUMEDREDGED ~ .,
                 data = gage_xgb_scaled,
                 method = "xgbTree",
                 tuneLength = 10,  # Automatic tuning of hyperparameters
                 trControl = ctrl
)

varImp(fit_dredge)
xgb.plot.tree(model = fit_dredge$finalModel, trees = 0)

fit_dredge

import_matrix = xgb.importance(model = fit_dredge$finalModel)
import_matrix
xgb.plot.importance(import_matrix)
```

```{r}
ggplot(data.frame(Observed = gage_xgb_scaled$VOLUMEDREDGED, Predicted = cv_pred), 
       aes(x = Observed, y = Predicted)) +
  geom_point() + 
  geom_abline(slope = 1, intercept = 0, color = "red") +
  labs(x = "Observed", y = "Predicted", title = "Observed vs Predicted")
```

### xgboost fast
```{r}
library(xgboost)
library(dplyr)

gage_survey_cleaned <- gage_survey_cleaned |>
  arrange(DATE_START)  # Sort the data by DATE_START

gage_xgb <- gage_survey_cleaned |>
  select(-lead_7, -lead_14,
         -DATE_START,
         -EXECYEAR,
         -RIVER, -Season) |>
  mutate_if(is.character, factor)

gage_xgb_scaled <- gage_xgb |>
  mutate(across(where(is.numeric), ~ scale(.) |> 
                  as.data.frame() |>
                  pull(1), .names = "{.col}"))

# 80/20 split for fast test
n <- nrow(gage_xgb_scaled)
split_idx <- floor(0.8 * n)

train_data <- gage_xgb_scaled[1:split_idx, ]
test_data  <- gage_xgb_scaled[(split_idx + 1):n, ]

train_data <- train_data %>%
  mutate(across(where(is.factor), ~ as.integer(.)))

test_data <- test_data %>%
  mutate(across(where(is.factor), ~ as.integer(.)))

# Create DMatrix
dtrain <- xgb.DMatrix(data = as.matrix(train_data %>% select(-VOLUMEDREDGED)),
                      label = train_data$VOLUMEDREDGED)
dtest <- xgb.DMatrix(data = as.matrix(test_data %>% select(-VOLUMEDREDGED)))

# Define quick params
params <- list(
  objective = "reg:squarederror",
  eta = 0.3,                # faster learning
  max_depth = 3,            # shallow trees for speed
  subsample = 0.8,
  colsample_bytree = 0.8
)

# Train with early stopping
model <- xgb.train(
  params = params,
  data = dtrain,
  nrounds = 50,             # keep small for fast testing
  watchlist = list(train = dtrain),
  early_stopping_rounds = 5,
  verbose = 0
)

# Predict and evaluate quickly
preds <- predict(model, dtest)
actuals <- test_data$VOLUMEDREDGED

# Quick RMSE
rmse <- sqrt(mean((actuals - preds)^2))
cat("Quick Test RMSE:", round(rmse, 3), "\n")

# Feature Importance Plot
import_matrix <- xgb.importance(model = model)
xgb.plot.importance(import_matrix)

```

###Cross Validation of Model
```{r}
library(progress)
pb <- progress_bar$new(total = n_outer_folds)
set.seed(123)
n = nrow(gage_xgb_scaled)
n_outer_folds = 5

cv_pred = vector(length = n)

groups = rep(1:n_outer_folds, length = n)
cv_groups = sample(groups, n)

for(ii in 1:n_outer_folds){
  print(Sys.time())
  pb$tick()
  in_test = (cv_groups == ii)
  in_train = (cv_groups != ii)

  ctrl = trainControl(method = "timeslice",
  initialWindow = 60,   # The first 60 data points for training
  horizon = 45,         # Forecasting horizon
  fixedWindow = TRUE,
  verboseIter=TRUE # Slide the window forward after each step
)
  fit_dredge_DCV = caret::train(VOLUMEDREDGED ~ .,
                       data = gage_xgb_scaled[in_train, ],
                       method = "xgbTree",
                     tune_grid = expand.grid(
  nrounds = c(25, 50),        # fewer rounds
  max_depth = c(2, 4),        # shallow trees
  eta = c(0.3),               # fast learning rate
  gamma = 0,                  # no regularization for speed
  colsample_bytree = 0.8,     # reasonable feature sampling
  min_child_weight = 1,
  subsample = 0.8
)
,
                       verbosity = 0,
                       trControl = ctrl,
                       na.action = na.exclude)

  # "Frankenstein" vector of predictions
  cv_pred[in_test] = predict(fit_dredge_DCV,
                             newdata = gage_xgb_scaled[in_test, ])
  print(Sys.time())

}

# Post-CV model evaluation

# RMSE: This is the most honest assessment of accuracy for regression
# I'll use RMSE to assess the model, because that's the metric I used to select the model.
# The model with the best RMSE may or may not be the one with the best MAE.
# RMSE
rmse = sqrt(mean((cv_pred - gage_xgb_scaled$VOLUMEDREDGED)^2))
cat("Root Mean Squared Error (RMSE):", rmse, "\n")

# Additional Regression Metrics (MAE, R-squared, etc.)
mae = mean(abs(cv_pred - gage_xgb_scaled$VOLUMEDREDGED))
cat("Mean Absolute Error (MAE):", mae, "\n")

# R-squared (R²) can be another good metric to evaluate model performance
ss_total = sum((gage_xgb_scaled$VOLUMEDREDGED - mean(gage_xgb_scaled$VOLUMEDREDGED))^2)
ss_residual = sum((gage_xgb_scaled$VOLUMEDREDGED - cv_pred)^2)
r_squared = 1 - (ss_residual / ss_total)
cat("R-squared:", r_squared, "\n")

# Best tuning parameters from last fold
best_tune <- fit_dredge_DCV$results
cat("Best Hyperparameters from Last Fold:\n")
print(best_tune)
```

## Well that didnt work! Lets try modeling between the two river systems

### Cleaning the data 
```{r}
combined_data <- bind_rows(survey_data, dredge_data)

combined_data <-combined_data |>
     filter(DATE_START >= as.Date("1999-04-01") &
         DATE_START <= as.Date("2024-12-31"))

combined_data <-combined_data |>
  mutate(lead_7 = (DATE_START - 7),
         lead_14 = (DATE_START - 14))
Nas<- combined_data|>
  filter(if_any(everything(),is.na))

gage_survey <- combined_data |>
  left_join(gage_data, by = "DATE_START")

gage_IWW <- gage_data |>
  select(DATE_START,BEAI2,HAVI2,
         HNYI2,IL03, IL03_Tail, IL04, IL04_Tail, IL05, IL05_Tail,IL06, IL06_Tail,
         IL07, IL07_Tail, IL08, IL08_Tail,KNGI2, MORI2, PIAI2,"05552500","05527500",
         "05585000","05583000","05570000","05555300")

gage_Miss <- gage_data |>
  select(DATE_START,BRLI4,CMMI4, DBQI4,
         FAII4, KHBI2,MI11,MI11_Tail,MI12,MI12_Tail,MI13,MI13_Tail,MI14,MI14_Tail,
         MI15,MI15_Tail,MI16,MI16_Tail,MI17,MI17_Tail,MI18,MI18_Tail,MI19,MI19_Tail,
         MI20,MI20_Tail,MI21,MI21_Tail,MI22,MI22_Tail, MUSI4,UINI2,"05446500")

Illinois_River <- gage_survey |>
  select(EXECYEAR,DATE_START,DREDGINGPURPOSE,VOLUMEDREDGED,RIVER,
         POOL, DREDGE_TYPE,UP_RIV_MIL, DN_RIV_MIL,lead_7,lead_14,BEAI2,HAVI2,
         HNYI2,IL03, IL03_Tail, IL04, IL04_Tail, IL05, IL05_Tail,IL06, IL06_Tail,
         IL07, IL07_Tail, IL08, IL08_Tail,KNGI2, MORI2, PIAI2,"05552500","05527500",
         "05585000","05583000","05570000","05555300") |>
  filter(RIVER == "Illinois_Waterway")

Miss_River <- gage_survey |>
  select(EXECYEAR,DATE_START,DREDGINGPURPOSE,VOLUMEDREDGED,RIVER,
         POOL, DREDGE_TYPE,UP_RIV_MIL, DN_RIV_MIL,lead_7,lead_14,BRLI4,CMMI4, DBQI4,
         FAII4, KHBI2,MI11,MI11_Tail,MI12,MI12_Tail,MI13,MI13_Tail,MI14,MI14_Tail,
         MI15,MI15_Tail,MI16,MI16_Tail,MI17,MI17_Tail,MI18,MI18_Tail,MI19,MI19_Tail,
         MI20,MI20_Tail,MI21,MI21_Tail,MI22,MI22_Tail, MUSI4,UINI2,"05446500") |>
  filter(RIVER == "Mississippi_River")

Miss_full <- Miss_River|>  
  left_join(gage_Miss, by = c("lead_7" = "DATE_START"), suffix = c("", "_7"))|>
  left_join(gage_Miss, by = c("lead_14" = "DATE_START"), suffix = c("","_14"))

IWW_full <- Illinois_River|>  
  left_join(gage_IWW, by = c("lead_7" = "DATE_START"), suffix = c("", "_7"))|>
  left_join(gage_IWW, by = c("lead_14" = "DATE_START"), suffix = c("","_14"))

Miss_Cleaned <- Miss_full |>
  filter(complete.cases(Miss_full))

IWW_Cleaned <- IWW_full |>
  filter(complete.cases(IWW_full))

IWW_Cleaned <- IWW_Cleaned |>
  mutate("Season" = sapply(DATE_START,date_to_season))

Miss_Cleaned <- Miss_Cleaned |>
  mutate("Season" = sapply(DATE_START,date_to_season))

```

## IWW PCA
```{r}
IWW_PCA <- IWW_Cleaned |>
select(where(is.numeric), -EXECYEAR)

IWW_prcomp <- IWW_PCA|>
  select(-VOLUMEDREDGED)

IWW_PCA = prcomp(IWW_prcomp, scale = T)

biplot(IWW_PCA)

plot(IWW_PCA,  type="l")

autoplot(IWW_PCA,data = IWW_Cleaned, 
         color = 'Season',loadings = TRUE,
         loadings.colour = 'black',
         loadings.label = TRUE, loadings.label.size = 4, 
         loadings.label.colour = "blue")

PCAvalues <- data.frame(Pool = IWW_Cleaned$POOL,
                        Dredge_Type = IWW_Cleaned$DREDGE_TYPE,
                        Season = IWW_Cleaned$Season,
                        Dredge_Occurance = IWW_Cleaned$DREDGINGPURPOSE,
                        IWW_PCA$x)

explained_variance <- (IWW_PCA$sdev^2) / sum(IWW_PCA$sdev^2) * 100

# Extract PC1 and PC2 explained variance
PC1_explained <- explained_variance[1]
PC2_explained <- explained_variance[2]

# Extract loadings of the variables
PCAloadings <- data.frame(Variables = rownames(IWW_PCA$rotation), IWW_PCA$rotation)

PCAloadings_PC1 <- PCAloadings|>
  arrange(desc(abs(PC1))) |>
  slice_head(n=10)

PCAloadings_PC2 <- PCAloadings|>
  arrange(desc(abs(PC2))) |>
  slice_head(n=10)

top_loadings <- bind_rows(PCAloadings_PC1, PCAloadings_PC2)

# Plot PCA with top 10 loadings
ggplot(PCAvalues, aes(x = PC1, y = PC2, colour = Season)) +
  geom_point(size = 2) +
  geom_segment(data = top_loadings, aes(x = 0, y = 0, xend = PC1 * 75, yend = PC2 * 75),
               arrow = arrow(length = unit(1/2, "picas")), color = "black") +  
  annotate("text", x = top_loadings$PC1 * 80, y = top_loadings$PC2 * 80,
           label = top_loadings$Variables, hjust = 0.2, vjust = 0.4) +  
  theme_minimal() +
  labs(title = "PCA Plot with Top 10 Loadings for PC1 and PC2",
      x = paste("PC1 (", round(PC1_explained, 1), "%)", sep = ""),
       y = paste("PC2 (", round(PC2_explained, 1), "%)", sep = ""))

ggplot(PCAvalues, aes(x = PC1, y = PC2, colour = Pool)) +
  geom_point(size = 2) +
  geom_segment(data = top_loadings, aes(x = 0, y = 0, xend = PC1 * 75, yend = PC2 * 75),
               arrow = arrow(length = unit(1/2, "picas")), color = "black") +  
  annotate("text", x = top_loadings$PC1 * 80, y = top_loadings$PC2 * 80,
           label = top_loadings$Variables, hjust = 0.2, vjust = 0.4) +  
  theme_minimal() +
  labs(title = "PCA Plot with Top 10 Loadings for PC1 and PC2",
      x = paste("PC1 (", round(PC1_explained, 1), "%)", sep = ""),
       y = paste("PC2 (", round(PC2_explained, 1), "%)", sep = ""))

ggplot(PCAvalues, aes(x = PC1, y = PC2, colour = Dredge_Type)) +
  geom_point(size = 2) + 
  geom_segment(data = top_loadings, aes(x = 0, y = 0, xend = PC1 * 75, yend = PC2 * 75),
               arrow = arrow(length = unit(1/2, "picas")), color = "black") + 
  annotate("text", x = top_loadings$PC1 * 80, y = top_loadings$PC2 * 80,
           label = top_loadings$Variables, hjust = 0.2, vjust = 0.4) +
  theme_minimal() +
  labs(title = "PCA Plot with Top 10 Loadings for PC1 and PC2",
      x = paste("PC1 (", round(PC1_explained, 1), "%)", sep = ""),
       y = paste("PC2 (", round(PC2_explained, 1), "%)", sep = ""))

ggplot(PCAvalues, aes(x = PC1, y = PC2, colour = Dredge_Occurance)) +
  geom_point(size = 2) + 
  geom_segment(data = top_loadings, aes(x = 0, y = 0, xend = PC1 * 75, yend = PC2 * 75),
               arrow = arrow(length = unit(1/2, "picas")), color = "black") +
  annotate("text", x = top_loadings$PC1 * 80, y = top_loadings$PC2 * 80,
           label = top_loadings$Variables, hjust = 0.2, vjust = 0.4) + 
  theme_minimal() +
  labs(title = "PCA Plot with Top 10 Loadings for PC1 and PC2",
      x = paste("PC1 (", round(PC1_explained, 1), "%)", sep = ""),
       y = paste("PC2 (", round(PC2_explained, 1), "%)", sep = ""))
```

## Miss PCA 
```{r}
Miss_PCA <- Miss_Cleaned |>
select(where(is.numeric), -EXECYEAR)

Miss_prcomp <- Miss_PCA|>
  select(-VOLUMEDREDGED)

Miss_PCA = prcomp(Miss_prcomp, scale = T)

biplot(Miss_PCA)
biplot(Miss_PCA, showLoadings = TRUE,)

plot(Miss_PCA,  type="l")

PCAvalues <- data.frame(Pool = Miss_Cleaned$POOL,
                        Dredge_Type = Miss_Cleaned$DREDGE_TYPE,
                        Season = Miss_Cleaned$Season,
                        Dredge_Occurance = Miss_Cleaned$DREDGINGPURPOSE,
                        Miss_PCA$x)

explained_variance <- (Miss_PCA$sdev^2) / sum(Miss_PCA$sdev^2) * 100

# Extract PC1 and PC2 explained variance
PC1_explained <- explained_variance[1]
PC2_explained <- explained_variance[2]

# Extract loadings of the variables
PCAloadings <- data.frame(Variables = rownames(Miss_PCA$rotation), Miss_PCA$rotation)

PCAloadings_PC1 <- PCAloadings|>
  arrange(desc(abs(PC1))) |>
  slice_head(n=10)

PCAloadings_PC2 <- PCAloadings|>
  arrange(desc(abs(PC2))) |>
  slice_head(n=10)

top_loadings <- bind_rows(PCAloadings_PC1, PCAloadings_PC2)

# Plot PCA with top 10 loadings
ggplot(PCAvalues, aes(x = PC1, y = PC2, colour = Season)) +
  geom_point(size = 2) +  # Scatter plot
  geom_segment(data = top_loadings, aes(x = 0, y = 0, xend = PC1 * 75, yend = PC2 * 75),
               arrow = arrow(length = unit(1/2, "picas")), color = "black") +  # Vectors for top loadings
  annotate("text", x = top_loadings$PC1 * 80, y = top_loadings$PC2 * 80,
           label = top_loadings$Variables, hjust = 0.2, vjust = 0.4) +  # Labels for top loadings
  theme_minimal() +
  labs(title = "PCA Plot with Top 10 Loadings for PC1 and PC2",
      x = paste("PC1 (", round(PC1_explained, 1), "%)", sep = ""),
       y = paste("PC2 (", round(PC2_explained, 1), "%)", sep = ""))

ggplot(PCAvalues, aes(x = PC1, y = PC2, colour = Pool)) +
  geom_point(size = 2) +  # Scatter plot
  geom_segment(data = top_loadings, aes(x = 0, y = 0, xend = PC1 * 75, yend = PC2 * 75),
               arrow = arrow(length = unit(1/2, "picas")), color = "black") +  # Vectors for top loadings
  annotate("text", x = top_loadings$PC1 * 80, y = top_loadings$PC2 * 80,
           label = top_loadings$Variables, hjust = 0.2, vjust = 0.4) +  # Labels for top loadings
  theme_minimal() +
  labs(title = "PCA Plot with Top 10 Loadings for PC1 and PC2",
      x = paste("PC1 (", round(PC1_explained, 1), "%)", sep = ""),
       y = paste("PC2 (", round(PC2_explained, 1), "%)", sep = ""))

ggplot(PCAvalues, aes(x = PC1, y = PC2, colour = Dredge_Type)) +
  geom_point(size = 2) +  # Scatter plot
  geom_segment(data = top_loadings, aes(x = 0, y = 0, xend = PC1 * 75, yend = PC2 * 75),
               arrow = arrow(length = unit(1/2, "picas")), color = "black") +  # Vectors for top loadings
  annotate("text", x = top_loadings$PC1 * 80, y = top_loadings$PC2 * 80,
           label = top_loadings$Variables, hjust = 0.2, vjust = 0.4) +  # Labels for top loadings
  theme_minimal() +
  labs(title = "PCA Plot with Top 10 Loadings for PC1 and PC2",
      x = paste("PC1 (", round(PC1_explained, 1), "%)", sep = ""),
       y = paste("PC2 (", round(PC2_explained, 1), "%)", sep = ""))

```

## xgboost split by river
#Miss
```{r}
Miss_xgb <-Miss_Cleaned |>
  select(#-lead_7,# -lead_14,
    -DATE_START,
         -EXECYEAR,-DREDGE_TYPE,
         -DREDGINGPURPOSE, -RIVER)|>
    mutate_if(is.character, factor)

ctrl = trainControl(method = "timeslice",
                      initialWindow = 60,   # The first 60 data points for training
                      horizon = 45,         # Forecasting horizon
                      fixedWindow = TRUE,   # Slide the window forward after each step
                      verboseIter = TRUE) 
Miss_xgb_fit = caret::train(VOLUMEDREDGED ~ .,
                 data = Miss_xgb,
                 method = "xgbTree",
               tuneGrid = expand.grid(nrounds = c(15,25,50,100), 
                                              max_depth = 1:10,
                                        eta =c(0.3), gamma = 0,
                                        colsample_bytree = 1,
                                        min_child_weight = 1, subsample = 1),
                 verbosity = 0,
                 trControl = ctrl,
                 na.action = na.exclude)

varImp(Miss_xgb_fit)
xgb.plot.tree(model = Miss_xgb_fit$finalModel, trees = 0)

Miss_xgb_fit

import_matrix = xgb.importance(model = Miss_xgb_fit$finalModel)
import_matrix
xgb.plot.importance(import_matrix)
```

#IWW
```{r}
IWW_xgb <-IWW_Cleaned |>
  select(-lead_7, -lead_14,-DATE_START,
         -EXECYEAR,-DREDGE_TYPE,
         -DREDGINGPURPOSE, -RIVER)|>
    mutate_if(is.character, factor)

ctrl = trainControl(method = "cv", number = 10) 
IWW_xgb_fit = caret::train(VOLUMEDREDGED ~ .,
                 data = IWW_xgb,
                 method = "xgbTree",
               tuneGrid = expand.grid(nrounds = c(15,25,50,100), 
                                              max_depth = 1:10,
                                        eta =c(0.3), gamma = 0,
                                        colsample_bytree = 1,
                                        min_child_weight = 1, subsample = 1),
                 verbosity = 0,
                 trControl = ctrl,
                 na.action = na.exclude)

varImp(IWW_xgb_fit)
xgb.plot.tree(model = IWW_xgb_fit$finalModel, trees = 0)

IWW_xgb_fit

import_matrix = xgb.importance(model = IWW_xgb_fit$finalModel)
import_matrix
xgb.plot.importance(import_matrix)
```
###LSTM
```{r}
library(tensorflow)
library(keras)

# Function to create sequences based on window size
create_sequences <- function(data, target_column, sequence_length) {
  X <- list()
  y <- list()

  for (i in (sequence_length + 1):nrow(data)) {
    X[[i - sequence_length]] <- as.matrix(data[(i - sequence_length):(i - 1), -target_column])  # Features excluding target
    y[[i - sequence_length]] <- data[i, target_column]  # Target column
  }

  X <- array(unlist(X), dim = c(length(X), sequence_length, ncol(data) - 1))  # Reshape to 3D array
  y <- unlist(y)

  return(list(X = X, y = y))
}

# Example: Create sequences for 7-day and 14-day windows
window_size_7 <- 7
window_size_14 <- 14

# Load necessary libraries
library(recipes)

# Assume gage_xgb is your original dataset
gage_xgb_scaled_recipe <- recipe(~ ., data = gage_xgb) |>
  step_dummy(all_nominal(), one_hot = TRUE) |>
  step_scale(all_numeric_predictors()) |>
  step_center(all_numeric_predictors())

# Prep the recipe
gage_xgb_prepped <- prep(gage_xgb_scaled_recipe, training = gage_xgb)

# Apply the scaling to the data
gage_xgb_scaled <- bake(gage_xgb_prepped, new_data = gage_xgb)

# View scaled data
head(gage_xgb_scaled)

gage_xgb_scaled_clean <- gage_xgb_scaled|>
  select(-lead_7,-lead_14)

# Assume `gage_xgb_scaled` is your scaled dataset and the target column is 'VOLUMEDREDGED'
data_windows_7 <- create_sequences(gage_xgb_scaled_clean, target_column = which(names(gage_xgb_scaled_clean) == "VOLUMEDREDGED"), sequence_length = window_size_7)
data_windows_14 <- create_sequences(gage_xgb_scaled_clean, target_column = which(names(gage_xgb_scaled_clean) == "VOLUMEDREDGED"), sequence_length = window_size_14)

# X_train and y_train for 7-day window
X_train_7 <- data_windows_7$X
y_train_7 <- data_windows_7$y

# X_train and y_train for 14-day window
X_train_14 <- data_windows_14$X
y_train_14 <- data_windows_14$y

# Define the train-test split
train_size_7 <- floor(0.8 * length(y_train_7))  # 80% for training
train_size_14 <- floor(0.8 * length(y_train_14))  # 80% for training

# Training and Testing Data for 7-day window
X_train_7_final <- X_train_7[1:train_size_7,,]
y_train_7_final <- y_train_7[1:train_size_7]
X_test_7_final <- X_train_7[(train_size_7+1):length(y_train_7),,]
y_test_7_final <- y_train_7[(train_size_7+1):length(y_train_7)]

# Training and Testing Data for 14-day window
X_train_14_final <- X_train_14[1:train_size_14,,]
y_train_14_final <- y_train_14[1:train_size_14]
X_test_14_final <- X_train_14[(train_size_14+1):length(y_train_14),,]
y_test_14_final <- y_train_14[(train_size_14+1):length(y_train_14)]

# Build the model
model_7 <- keras_model_sequential() |>
  layer_lstm(units = 50, return_sequences = TRUE, input_shape = c(7, ncol(gage_xgb_scaled_clean) - 1))  |>
  layer_dropout(rate = 0.2)|>
  layer_lstm(units = 50, return_sequences = TRUE) |>
  layer_dropout(rate = 0.2) |>
  layer_lstm(units = 50) |>
  layer_dropout(rate = 0.2) |>
  layer_dense(units = 1)

# model_7 %>% compile(
#   loss = 'mean_squared_error',
#   optimizer = 'adam'
# )

model_7 %>% compile(
  loss = 'mean_squared_error',
  optimizer = optimizer_adam(learning_rate = 0.001)  # Try also 0.0005 or 0.0001
)

callback <- callback_early_stopping(monitor = "val_loss", patience = 10, restore_best_weights = TRUE)

history_7 <- model_7 |> fit(
  X_train_7_final, 
  y_train_7_final,
  epochs = 1000,
  batch_size = 32,
  validation_data = list(X_test_7_final, y_test_7_final),
  callbacks = list(callback)
)

# Train the model for the 7-day window
# history_7 <- model_7 %>% fit(
#   X_train_7_final, 
#   y_train_7_final,
#   epochs = 1000, 
#   batch_size = 32, 
#   validation_data = list(X_test_7_final, y_test_7_final)
# )

# Evaluate model on 7-day window
score_7 <- model_7 %>% evaluate(X_test_7_final, y_test_7_final)
cat('Test loss for 7-day window:', score_7, "\n")

predictions_7 <- model_7 %>% predict(X_test_7_final)

```

### LSTM method2
```{r}
library(tensorflow)
library(keras)
library(recipes)
library(dplyr)

# Set window sizes
window_size_7 <- 7
window_size_14 <- 14

# --- Step 1: Separate and scale features & target ---
target_col <- "VOLUMEDREDGED"
gage_target <- gage_xgb[[target_col]]
gage_features <- gage_xgb |> select(-all_of(target_col))

# Scale predictors
gage_recipe <- recipe(~ ., data = gage_features) |>
  step_dummy(all_nominal(), one_hot = TRUE) |>
  step_scale(all_numeric_predictors()) |>
  step_center(all_numeric_predictors())
gage_recipe_prep <- prep(gage_recipe, training = gage_features)
gage_features_scaled <- bake(gage_recipe_prep, new_data = gage_features)

# Scale target manually
target_mean <- mean(gage_target, na.rm = TRUE)
target_sd <- sd(gage_target, na.rm = TRUE)
gage_target_scaled <- (gage_target - target_mean) / target_sd

# Recombine scaled features and target
gage_scaled_clean <- bind_cols(gage_features_scaled, VOLUMEDREDGED = gage_target_scaled)

# Optionally remove future-looking targets
gage_scaled_clean <- gage_scaled_clean |> select(-lead_7, -lead_14)

# --- Step 2: Sequence creation function ---
create_sequences <- function(data, target_column, sequence_length) {
  X <- list()
  y <- list()

  for (i in (sequence_length + 1):nrow(data)) {
    X[[i - sequence_length]] <- as.matrix(data[(i - sequence_length):(i - 1), -target_column])
    y[[i - sequence_length]] <- data[i, target_column]
  }

  X <- array(unlist(X), dim = c(length(X), sequence_length, ncol(data) - 1))
  y <- unlist(y)

  return(list(X = X, y = y))
}

# --- Step 3: Create sequences ---
target_index <- which(names(gage_scaled_clean) == "VOLUMEDREDGED")
data_windows_7 <- create_sequences(gage_scaled_clean, target_column = target_index, sequence_length = window_size_7)

X <- data_windows_7$X
y <- data_windows_7$y

# Split into train/test
train_size <- floor(0.8 * length(y))
X_train <- X[1:train_size,,]
y_train <- y[1:train_size]
X_test <- X[(train_size+1):length(y),,]
y_test <- y[(train_size+1):length(y)]

# --- Step 4: Build model ---
model <- keras_model_sequential() |>
  layer_lstm(units = 50, return_sequences = TRUE, input_shape = c(window_size_7, ncol(gage_scaled_clean) - 1)) |>
  layer_dropout(rate = 0.2) |>
  layer_lstm(units = 50, return_sequences = TRUE) |>
  layer_dropout(rate = 0.2) |>
  layer_lstm(units = 50) |>
  layer_dropout(rate = 0.2) |>
  layer_dense(units = 1)

model %>% compile(
  loss = 'mean_squared_error',
  optimizer = optimizer_adam(learning_rate = 0.001)
)

callback <- callback_early_stopping(monitor = "val_loss", patience = 10, restore_best_weights = TRUE)

history <- model %>% fit(
  X_train, y_train,
  epochs = 1000,
  batch_size = 32,
  validation_data = list(X_test, y_test),
  callbacks = list(callback)
)

# --- Step 5: Predict and invert scaling ---
pred_scaled <- model %>% predict(X_test)
pred <- pred_scaled * target_sd + target_mean
actual <- y_test * target_sd + target_mean

# --- Step 6: Evaluate on original scale ---
rmse <- sqrt(mean((actual - pred)^2))
r_squared <- 1 - sum((actual - pred)^2) / sum((actual - mean(actual))^2)
mae <- mean(abs(actual - pred))

cat("RMSE:", rmse, "\n")
cat("R-squared:", r_squared, "\n")
cat("MAE:", mae, "\n")
```

##Lets try predicting a binary yes/no for dredging
```{r}
library(tensorflow)
library(keras)
library(recipes)
library(dplyr)

# --- Step 1: Parameters ---
window_size <- 7
forecast_horizon <- 7  # Predict if dredging will occur in the next 7 days
target_col <- "VOLUMEDREDGED"

# --- Step 2: Feature/target separation and scaling ---
gage_target <- gage_xgb[[target_col]]
gage_features <- gage_xgb |> select(-all_of(target_col))

# Use recipes to scale and encode predictors
gage_recipe <- recipe(~ ., data = gage_features) |>
  step_dummy(all_nominal(), one_hot = TRUE) |>
  step_scale(all_numeric_predictors()) |>
  step_center(all_numeric_predictors())

gage_recipe_prep <- prep(gage_recipe, training = gage_features)
gage_features_scaled <- bake(gage_recipe_prep, new_data = gage_features)

# Keep target as-is (not scaled), used for binary labeling
gage_scaled_clean <- bind_cols(gage_features_scaled, VOLUMEDREDGED = gage_target)

# Optional: remove future-looking columns
gage_scaled_clean <- gage_scaled_clean |> select(-lead_7, -lead_14)

# --- Step 3: Sequence creation for classification ---
create_binary_sequences <- function(data, target_column, sequence_length, forecast_horizon) {
  X <- list()
  y <- list()

  for (i in (sequence_length + 1):(nrow(data) - forecast_horizon)) {
    X[[i - sequence_length]] <- as.matrix(data[(i - sequence_length):(i - 1), -target_column])

    future_window <- data[i:(i + forecast_horizon - 1), target_column]
    dredging_needed <- as.integer(any(future_window > 0))

    y[[i - sequence_length]] <- dredging_needed
  }

  X <- array(unlist(X), dim = c(length(X), sequence_length, ncol(data) - 1))
  y <- unlist(y)

  return(list(X = X, y = y))
}

# --- Step 4: Generate data ---
target_index <- which(names(gage_scaled_clean) == "VOLUMEDREDGED")
data_windows <- create_binary_sequences(
  gage_scaled_clean,
  target_column = target_index,
  sequence_length = window_size,
  forecast_horizon = forecast_horizon
)

X <- data_windows$X
y <- data_windows$y

# --- Step 5: Train/test split ---
train_size <- floor(0.8 * length(y))
X_train <- X[1:train_size,,]
y_train <- y[1:train_size]
X_test <- X[(train_size + 1):length(y),,]
y_test <- y[(train_size + 1):length(y)]

# --- Step 6: LSTM model ---
model <- keras_model_sequential() |>
  layer_lstm(units = 50, return_sequences = TRUE, input_shape = c(window_size, dim(X)[3])) |>
  layer_dropout(rate = 0.2) |>
  layer_lstm(units = 50, return_sequences = TRUE) |>
  layer_dropout(rate = 0.2) |>
  layer_lstm(units = 50) |>
  layer_dropout(rate = 0.2) |>
  layer_dense(units = 1, activation = "sigmoid")  # Binary output

model %>% compile(
  loss = 'binary_crossentropy',
  optimizer = optimizer_adam(learning_rate = 0.001),
  metrics = c('accuracy', metric_auc())
)

# --- Step 7: Train model ---
callback <- callback_early_stopping(monitor = "val_loss", patience = 10, restore_best_weights = TRUE)

history <- model %>% fit(
  X_train, y_train,
  epochs = 100,
  batch_size = 32,
  validation_data = list(X_test, y_test),
  callbacks = list(callback)
)

# --- Step 8: Evaluate and predict ---
pred_probs <- model %>% predict(X_test)
pred_classes <- ifelse(pred_probs > 0.5, 1, 0)

# Confusion matrix
confusion_matrix <- table(Predicted = pred_classes, Actual = y_test)
print(confusion_matrix)

# Metrics
accuracy <- mean(pred_classes == y_test)
TP <- sum(pred_classes == 1 & y_test == 1)
FP <- sum(pred_classes == 1 & y_test == 0)
FN <- sum(pred_classes == 0 & y_test == 1)

precision <- ifelse((TP + FP) > 0, TP / (TP + FP), 0)
recall <- ifelse((TP + FN) > 0, TP / (TP + FN), 0)
f1 <- ifelse((precision + recall) > 0, 2 * precision * recall / (precision + recall), 0)

cat("Accuracy:", accuracy, "\n")
cat("Precision:", precision)

```

```{r}
library(tensorflow)
library(keras)
library(recipes)
library(dplyr)

prepare_lstm_data <- function(data) {
  # Select relevant features
  relevant_features <- c(
    "VOLUMEDREDGED", "UP_RIV_MIL", "DN_RIV_MIL", "POOL", "DATE_START",
    names(select(data, matches("MI\\d+")))
  )

  data_subset <- data %>%
    select(all_of(relevant_features)) %>%
    arrange(DATE_START)

  # Convert POOL to factor and then numeric
  data_subset$POOL <- as.numeric(factor(data_subset$POOL))

  # Create binary target
  data_subset$dredge_needed <- as.integer(data_subset$VOLUMEDREDGED > 0)

  # Remove DATE_START after arranging
  data_subset <- data_subset %>%
    select(-DATE_START)

  # Scale numeric features
  recipe_obj <- recipe(~ ., data = data_subset) %>%
    step_normalize(all_numeric_predictors())

  prepped_recipe <- prep(recipe_obj)
  scaled_data <- bake(prepped_recipe, new_data = data_subset)

  return(scaled_data)
}

# 2. Modified sequence creation
create_sequences <- function(data, window_size = 45) {
  # Ensure all columns are numeric
  numeric_data <- data %>%
    mutate(across(everything(), as.numeric))

  X <- list()
  y <- list()

  for(i in 1:(nrow(numeric_data) - window_size)) {
    # Exclude dredge_needed column from features
    X[[i]] <- as.matrix(numeric_data[i:(i + window_size - 1), 
                                   !names(numeric_data) %in% c("dredge_needed")])
    y[[i]] <- numeric_data$dredge_needed[i + window_size]
  }

  X <- array(unlist(X), 
            dim = c(length(X), 
                   window_size, 
                   ncol(numeric_data) - 1))  # -1 for dredge_needed column
  y <- unlist(y)

  return(list(X = X, y = y))
}

# 3. Main execution
# Prepare data
scaled_data <- prepare_lstm_data(Miss_Cleaned)

# Create sequences
sequences <- create_sequences(scaled_data)

# Split data
split_idx <- floor(0.8 * length(sequences$y))
X_train <- sequences$X[1:split_idx,,]
y_train <- sequences$y[1:split_idx]
X_val <- sequences$X[(split_idx+1):length(sequences$y),,]
y_val <- sequences$y[(split_idx+1):length(sequences$y)]

# Verify data types
print("Training data dimensions:")
print(dim(X_train))
print("Training labels length:")
print(length(y_train))
print("First few values of X_train:")
print(head(X_train[1,,]))
print("First few values of y_train:")
print(head(y_train))

# Train model
model_results <- train_lstm_model(X_train, y_train, X_val, y_val)

# Evaluate model
predictions <- predict(model_results$model, X_val)
binary_predictions <- ifelse(predictions > 0.5, 1, 0)

# Calculate metrics
confusion_matrix <- table(Actual = y_val, Predicted = binary_predictions)
accuracy <- sum(diag(confusion_matrix)) / sum(confusion_matrix)
recall <- confusion_matrix[2,2] / sum(confusion_matrix[2,])

print("Confusion Matrix:")
print(confusion_matrix)
print(paste("Accuracy:", round(accuracy, 3)))
print(paste("Recall:", round(recall, 3)))

# First, let's check the distribution of actual values
print("Distribution of actual values (y_val):")
print(table(y_val))

# Check the raw predictions before binary conversion
print("Summary of raw predictions:")
print(summary(predictions))

# Modified evaluation code with error handling
evaluate_binary_predictions <- function(y_true, y_pred) {
  # Convert predictions to binary
  binary_pred <- ifelse(y_pred > 0.5, 1, 0)

  # Create confusion matrix
  conf_matrix <- table(Actual = y_true, Predicted = binary_pred)
  print("Confusion Matrix:")
  print(conf_matrix)

  # Calculate metrics with error handling
  accuracy <- sum(diag(conf_matrix)) / sum(conf_matrix)

  # Handle case where there are no positive predictions
  if(length(conf_matrix) == 1 || !any(binary_pred == 1)) {
    recall <- 0
    precision <- 0
    f1 <- 0
  } else {
    TP <- conf_matrix["1", "1"]
    FN <- conf_matrix["1", "0"]
    FP <- conf_matrix["0", "1"]

    recall <- ifelse(TP + FN > 0, TP / (TP + FN), 0)
    precision <- ifelse(TP + FP > 0, TP / (TP + FP), 0)
    f1 <- ifelse(precision + recall > 0, 
                 2 * (precision * recall) / (precision + recall), 
                 0)
  }

  return(list(
    confusion_matrix = conf_matrix,
    accuracy = accuracy,
    recall = recall,
    precision = precision,
    f1 = f1
  ))
}

# Modify the model to address class imbalance
train_lstm_model <- function(X_train, y_train, X_val, y_val) {
  # Calculate class weights
  n_neg <- sum(y_train == 0)
  n_pos <- sum(y_train == 1)
  total <- n_neg + n_pos

  class_weight <- list(
    "0" = 1,
    "1" = n_neg/n_pos  # Give more weight to minority class
  )

  model <- keras_model_sequential() %>%
    layer_lstm(units = 64, return_sequences = TRUE, 
              input_shape = c(dim(X_train)[2], dim(X_train)[3])) %>%
    layer_dropout(0.2) %>%
    layer_lstm(units = 32) %>%
    layer_dropout(0.2) %>%
    layer_dense(units = 16, activation = "relu") %>%
    layer_dense(units = 1, activation = "sigmoid")

  model %>% compile(
    loss = "binary_crossentropy",
    optimizer = optimizer_adam(learning_rate = 0.001),
    metrics = c("accuracy", "AUC")
  )

  history <- model %>% fit(
    X_train, y_train,
    epochs = 50,
    batch_size = 32,
    validation_data = list(X_val, y_val),
    class_weight = class_weight,  # Add class weights
    callbacks = list(
      callback_early_stopping(
        monitor = "val_loss",
        patience = 10,
        restore_best_weights = TRUE
      )
    )
  )

  return(list(model = model, history = history))
}

# Retrain model with class weights
model_results <- train_lstm_model(X_train, y_train, X_val, y_val)

# Make predictions
predictions <- predict(model_results$model, X_val)

# Evaluate with new function
results <- evaluate_binary_predictions(y_val, predictions)

# Print detailed results
print("Evaluation Results:")
print(paste("Accuracy:", round(results$accuracy, 3)))
print(paste("Recall:", round(results$recall, 3)))
print(paste("Precision:", round(results$precision, 3)))
print(paste("F1 Score:", round(results$f1, 3)))

# Plot training history
plot(model_results$history)

# Additional analysis of predictions
hist(predictions, 
     main = "Distribution of Predictions",
     xlab = "Predicted Probability",
     breaks = 20)

```