### 2025-06-26 CSAT_Rates.Rmd
CSAT rate wrangling and spot checks

---
title: "CSAT_Wrangling_ShoalingRates"
output: html_document
date: "2025-05-12"
---

```{r}
library(readr)
library(dplyr)
library(stringr)
library(purrr)
library(ggplot2)
library(lubridate)
library(tidyverse)
library(zoo)
```

```{r}
# Set folder path
folder_path <- "C:/workspace/CSAT/CSAT_distribution/output/CEMVR/19990101__to__20241231"

reach_data<-read_csv("C:/workspace/DRAFT/ReachData.csv")

```

```{r, message=FALSE}
### Shoaling Rates

files <-list.files(path = folder_path,
                                 pattern = "_SurveyPairVolumeDifference\\.csv$",
                                 full.names = TRUE)

read_and_label2 <- function(file_path) {
  # Extract filename
  file_name <- basename(file_path)

  # Remove extension
  file_name_no_ext <- tools::file_path_sans_ext(file_name)

  # Split by underscore
  parts <- str_split(file_name_no_ext, "_", simplify = TRUE)

  # Extract relevant parts
  river <- parts[1]
  pool <- parts[2]
  reach <- parts[4]  # MVR is at [3], 14 or 22 is at [4]

  # Read the data
  data <- read_csv(file_path, show_col_types = FALSE)

  # Add new columns
  data <- data %>%
    mutate(
      river = river,
      pool = pool,
      reach = reach,
      source_file = file_name
    )

  return(data)
}

# Apply the function to all files and bind into one big dataframe
combined_data <- map_dfr(files, read_and_label2)
```

```{r}
# Check it out
glimpse(combined_data)

combined_data_clean<-combined_data|>
  mutate(Before_date = ymd(SurveyDateBefore),
         After_date = ymd(SurveyDateAfter))|>
  mutate(DateDiff = as.numeric(After_date - Before_date))|>
  mutate(DailyShoalingRate_ftperday = AnnualShoalingRate_ftperyr/DateDiff,
         DailyShoalingVolume_Cyperday = AnnualShoalingVolume_CYperyr/DateDiff,
         VolChange_Cyperday = NetVolumeChange_CY/DateDiff)
#  filter(DailyShoalingVolume_Cyperday < 12500000 & DailyShoalingRate_ftperday < 900)
```

###Plot Rates by day
```{r}
ggplot(combined_data_clean, aes(x = After_date, y = AnnualShoalingVolume_CYperyr, colour = river)) +
  geom_point() + 
  geom_line()+
#  ylim(-2000000,1000000)+
  facet_wrap(~ river)

ggplot(combined_data_clean, aes(x = After_date, y = AnnualShoalingRate_ftperyr, colour = river)) +
  geom_point() + 
  geom_line()+
#  ylim(NA, 275)+
  facet_wrap(~ river)

ggplot(combined_data_clean, aes(x = After_date, y = NetVolumeChange_CY, colour = river)) +
  geom_point() + 
  geom_line()+
#  ylim(NA, 275)+
  facet_wrap(~ river)

ggplot(combined_data_clean, aes(x = Before_date, y = NetVolumeChange_CY)) +
  geom_point() +
  theme(legend.position = "bottom")+
  geom_line(alpha = 0.5) +
  labs(title = "Net Volume Change in Sediment (CY) Over Time",
       x = "Date of Survey After",
       y = "Net Volume Change (CY)",
       color = "River") +
  theme_minimal() +
    facet_wrap(~ pool, scales = "free_y")

```
Let's only look at rate differences calculated within a year 
```{r}
combined_data_year <- combined_data_clean|>
  mutate(yr_diff = abs(year(After_date) - year(Before_date)))|>
  filter(yr_diff <1)|>
  filter(pool != "AL")|>
  filter(pool != "CS")
```

```{r}
ggplot(combined_data_year, aes(x = After_date, y = AnnualShoalingVolume_CYperyr, colour = river)) +
  geom_point() + 
  geom_line()+
#  ylim(NA,2000000)+
  facet_wrap(~ river)

ggplot(combined_data_year, aes(x = After_date, y = AnnualShoalingRate_ftperyr, colour = river)) +
  geom_point() + 
  geom_line()+
#  ylim(NA, 275)+
  facet_wrap(~ river)

ggplot(combined_data_year, aes(x = After_date, y = NetVolumeChange_CY, colour = river)) +
  geom_point() + 
  geom_line()+
#  ylim(NA, 275)+
  facet_wrap(~ river)

ggplot(combined_data_year, aes(x = Before_date, y = NetVolumeChange_CY, color = river)) +
  geom_point() +
  theme(legend.position = "bottom")+
  geom_line(alpha = 0.5) +
  labs(title = "Net Volume Change in Sediment (CY) Over Time By Pool",
       x = "Survey Date",
       y = "Net Volume Change (CY)",
       color = "River") +
  theme_minimal() +
    facet_wrap(~ pool, scales = "free_y")
```

### different view
```{r}
daily_expanded <- combined_data_clean |>
  # Create a sequence of dates for each row
  rowwise() |>
  mutate(DateSeq = list(seq(Before_date, After_date, by = "day"))) |>
  ungroup() |>
  # Expand to one row per day
  unnest(cols = c(DateSeq)) |>
  rename(Date = DateSeq) |>
  # Keep only the relevant daily rate values
  select(Date, DailyShoalingVolume_Cyperday,DailyShoalingRate_ftperday, river, pool, reach)

write_csv(daily_expanded, "C:/workspace/DRAFT/CSAT_Data/SurveyPairVolumeDifference_PreRollingAvg.csv")

  glimpse(daily_expanded)

# daily_series <- daily_expanded |>
#   group_by(Date, river, pool, reach) |>
#   summarise(DailyVolume = sum(DailyShoalingVolume_Cyperday, na.rm = TRUE), 
#             DailyRate = sum(DailyShoalingRate_ftperday, na.rm = TRUE),
#             .groups = "drop") |>
#   filter(pool == "LA")|>
#   arrange(river, pool, reach, Date) |>
#   group_by(river, pool, reach) |>
#   summarise(mutate(RollingAvgVol_7d = rollmean(DailyVolume, k = 7, fill = NA, align = "right"),
#          RollingAvgRate_7d = rollmean(DailyRate, k = 7, fill = NA, align = "right")))  |>
#   ungroup()

daily_avg <- daily_expanded %>%
  group_by(Date, river, pool, reach) %>%
  summarise(
    avg_volume = mean(DailyShoalingVolume_Cyperday, na.rm = TRUE),
    count_volume = sum(!is.na(DailyShoalingVolume_Cyperday)),
     sd_volume = sd(DailyShoalingVolume_Cyperday, na.rm = TRUE),
    se_volume = sd_volume/sqrt(n()),
    avg_rate = mean(DailyShoalingRate_ftperday, na.rm = TRUE),
    count_rate = sum(!is.na(DailyShoalingRate_ftperday)),
    sd_rate = sd(DailyShoalingRate_Cyperday, na.rm = TRUE),
    se_rate = sd_rate/sqrt(n()),
    .groups = "drop"
  )|>
  mutate(site = paste(river, pool, reach, sep = "_"))

daily_avg_rolled <- daily_avg %>%
  arrange(site, Date) %>%
  group_by(site) %>%
  mutate(
    avg_volume_7day = rollmean(avg_volume, k = 7, fill = NA, align = "right"),
     avg_volume_7day_upper = avg_volume_7day + 1.96 * rollmean(se_volume, k = 7, fill = NA),
    avg_volume_7day_lower = avg_volume_7day - 1.96 * rollmean(se_volume, k = 7, fill = NA),
    avg_rate_7day = rollmean(avg_rate, k = 7, fill = NA, align = "right"),
  ) %>%
  ungroup()

write_csv(daily_avg_rolled, "C:/workspace/DRAFT/CSAT_Data/SurveyPairVolumeDifference.csv")

# Loop through each unique value in the `pool` column
for (pool_value in unique(daily_avg_rolled$pool)) {

# Create a new dataframe for each pool value, named dynamically
  assign(paste0("P_", pool_value), daily_avg_rolled %>% filter(pool == pool_value))
}

year_plots<-year_plots |>
  mutate(Date_norm = as.Date("2000-01-01") + as.numeric(Date - lubridate::floor_date(Date, "year")))

year_plots |>
  ggplot(aes(Date_norm, avg_rate_7day, color = factor(year))) +
  geom_line() +
  scale_x_date(date_labels = "%b")+
  facet_wrap(~reach,scales = "free_y")+
  ggtitle("Pool 18 avg 7 day volumes across each reach")

daily_avg_rolled|>
  filter(river == "IL")|>
  ggplot(aes(x = Date, y = avg_volume_7day, colour = river)) +
  geom_point() + 
  ylim(NA,125000)+
  geom_line()

daily_avg_rolled|>
  filter(river == "UM")|>
  ggplot(aes(x = Date, y = avg_volume_7day, colour = river)) +
  geom_point() + 
  ylim(NA,125000)+
  geom_line()

daily_avg_rolled |>
  filter(river == "IL")|>
  ggplot(aes(x = Date, y = count_rate)) +
  geom_col(fill = "dodgerblue", alpha = 0.60)+
  facet_wrap(~pool)

daily_avg_rolled |>
  filter(river == "UM")|>
  ggplot(aes(x = Date, y = count_rate)) +
  geom_col(fill = "dodgerblue", alpha = 0.60)+
  facet_wrap(~pool)

#, scales = "free_y"

# Then: pivot to wide format
daily_volume_wide <- daily_avg %>%
  select(Date, site, avg_volume) %>%
  pivot_wider(
    names_from = site,
    values_from = avg_volume
  )|>
  mutate(year = year(Date))

daily_rate_wide <- daily_avg %>%
  select(Date, site, avg_rate) %>%
  pivot_wider(
    names_from = site,
    values_from = avg_rate
  )

daily_expanded |>
  filter(pool == "20")|>
  filter(year(Date) >= 2015 & year(Date) <= 2020)|>
  ggplot(aes(x = Date, y = DailyShoalingVolume_Cyperday, colour = river)) +
  geom_point() + 
  geom_line()+
  facet_wrap(~reach)

```

```{r}
# For the yearly comparison plot
year_plots |>
  ggplot(aes(Date_norm, avg_rate_7day, color = factor(year))) +
  geom_line() +
  scale_x_date(date_labels = "%b") +
  geom_ribbon(aes(ymin = avg_rate_7day_lower, ymax = avg_rate_7day_upper, fill = factor(year)), alpha = 0.2) +
  facet_wrap(~reach, scales = "free_y") +
  theme_minimal() +
  labs(
    title = "Pool 18 Average 7-day Volumes Across Each Reach",
    x = "Month",
    y = "7-day Average Rate",
    color = "Year"
  )

# For the river-specific plots
daily_avg_rolled |>
  filter(river == "IL") |>
  ggplot(aes(x = Date, y = avg_volume_7day)) +
  geom_line(color = "blue") +
   geom_ribbon(aes(ymin = avg_volume_7day_lower, ymax = avg_volume_7day_upper), 
              fill = "blue", alpha = 0.2) +
  ylim(NA, 125000) +
  theme_minimal() +
  labs(
    title = "Illinois River 7-day Average Volumes",
    x = "Date",
    y = "Volume (CY)"
  )
```