### 2025-09-03 Lesson 6 examples--XGBoost.Rmd

---
title: "XGBoost"
author: "Abra Brisbin"
date: "11/1/2020"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(ggformula)

library(ISLR) # for College data, used in hwk 4

library(caret)

library(xgboost)
library(DiagrammeR)

```

## Predicting Enrollment

```{r}

head(College)

```

```{r}

set.seed(123)
ctrl = trainControl(method = "cv", number = 5) # Increase this to 10 to get more consistent RMSEs
fit_coll = train(Enroll ~ ., 
             data = College,
             method = "xgbTree",
             tuneGrid = expand.grid(nrounds = 100, max_depth = 1:3, 
                                    eta = 0.3, gamma = 0, 
                                    colsample_bytree = .8, 
                                    min_child_weight = 1, subsample = 1),
             verbosity = 0, 
             trControl = ctrl)

fit_coll

```

### Variable importance

```{r}

varImp(fit_coll)

```

```{r}

import_matrix = xgb.importance(model = fit_coll$finalModel)
import_matrix
xgb.plot.importance(import_matrix)

```
## Double CV

```{r}

set.seed(123)
n = nrow(College)
n_outer_folds = 10

cv_pred = vector(length = n)

groups = rep(1:n_outer_folds, length = n)
cv_groups = sample(groups, n)

for(ii in 1:n_outer_folds){
  in_test = (cv_groups == ii)
  in_train = (cv_groups != ii)

  ctrl = trainControl(method = "cv", number = 5)
  fit_coll_DCV = train(Enroll ~ ., 
                       data = College[in_train, ],
                       method = "xgbTree",
                       tuneGrid = expand.grid(nrounds = 100, max_depth = 1:3, 
                                              eta = 0.3, gamma = 0, 
                                              colsample_bytree = .8, 
                                              min_child_weight = 1, subsample = 1),
                       verbosity = 0, 
                       trControl = ctrl)

  # "Frankenstein" vector of predictions
  cv_pred[in_test] = predict(fit_coll_DCV, 
                             newdata = College[in_test, ])

}

# I'll use RMSE to assess the model, because that's the metric I used to select the model.
# The model with the best RMSE may or may not be the one with the best MAE.
# RMSE
sqrt(mean((cv_pred - College$Enroll)^2)) # This is the most honest 
                                         # assessment of accuracy

```

## Viewing Trees

Let's make a model with 2 trees, because that will be easier to view and make predictions about.
```{r}

set.seed(123)
ctrl = trainControl(method = "cv", number = 5)
fit_simple2 = train(Enroll ~ ., 
             data = College,
             method = "xgbTree",
             tuneGrid = expand.grid(nrounds = 2, max_depth = 1:2, 
                                    eta = 0.3, gamma = 0, 
                                    colsample_bytree = .8, 
                                    min_child_weight = 1, 
                                    subsample = 1),
             verbosity = 0, 
             trControl = ctrl)

```

```{r}

xgb.plot.tree(model = fit_simple2$finalModel, trees = 0)
xgb.plot.tree(model = fit_simple2$finalModel, trees = 1)
```

## Relating predictions to the trees
```{r}
# For a regression tree, the predictions are simply the sum of the Values in their leaves, + .5.

predictions2 = predict(fit_simple2)

head(predictions2)
College %>%
  dplyr::select(F.Undergrad, Accept) %>%
  head()
```
Abilene Christian University:  Tree 0 value = 100.426697
Tree 1 value = 202.066833
```{r}
100.426697 + 202.066833 +.5 # Matches the predicted value from predict()
```
### Classification trees

```{r}
set.seed(123)
ctrl = trainControl(method = "cv", number = 5)
private_simple2 = train(Private ~ ., 
             data = College,
             method = "xgbTree",
             tuneGrid = expand.grid(nrounds = 2, max_depth = 1:2, eta = 0.3, gamma = 0, colsample_bytree = .8, min_child_weight = 1, subsample = 1),
             trControl = ctrl)

private_simple2
```

```{r}
xgb.plot.tree(model = private_simple2$finalModel, trees = 0)
```

```{r}
xgb.plot.tree(model = private_simple2$finalModel, trees = 1)
```

```{r}
predictions_priv = predict(private_simple2, type = "prob")
head(predictions_priv)
College %>%
  dplyr::select(Outstate, Accept, Enroll, F.Undergrad, Room.Board) %>%
  head()
```
Abilene Christian University:  Tree 0: -0.156000003
Tree 1:  -0.412369579
```{r}
# For classification, the sum of the Values in the relevant leaves
# gives the log odds of Y = 1
a = 0 + exp(-0.156000003-0.412369579)
1/(1+a) # Prob(Private), matches the value from predict()
```

Y = 1 is "Yes", Private.
```{r}
contrasts(College$Private)
```

##################################

### 1 Tree
```{r}

set.seed(123)
ctrl = trainControl(method = "cv", number = 5)
fit_simple = train(Enroll ~ ., 
             data = College,
             method = "xgbTree",
             tuneGrid = expand.grid(nrounds = 1, max_depth = 1, 
                                    eta = 0.3, gamma = 0, 
                                    colsample_bytree = .8, 
                                    min_child_weight = 1, 
                                    subsample = 1),
             verbosity = 0, # suppresses a warning about 
                            # deprecated version of the 
                            # objective function
                            # (not necessary for classification)
             trControl = ctrl)

fit_simple

```

### 100 Trees

```{r}

#xgb.plot.tree(model = fit_coll$finalModel) # would show all trees
xgb.plot.tree(model = fit_coll$finalModel, trees = 0)
```

### Regression tree predictions
```{r}
xgb.plot.tree(model = fit_simple$finalModel)
```
```{r}
# baseline = 0.5, so each prediction is 0.5 + the Value of its leaf.
# darker arrows are the "yes" direction.
predictions = predict(fit_simple)
head(predictions)
head(College$Accept)
```

```{r}

xgb.plot.tree(model = fit_simple2$finalModel, trees = 0)

```

```{r}
xgb.plot.tree(model = fit_simple2$finalModel, trees = 1)
```

```{r}
# For a regression tree, the predictions are simply the sum of the Values in their leaves, + .5.

predictions2 = predict(fit_simple2)

head(predictions2)
College %>%
  select(F.Undergrad, Apps) %>%
  head()
```
Abilene Christian University:  Tree 0 value = 221.183502
Tree 1 value = 101.9478
```{r}
221.183502 + 101.9478 +.5 # Matches the predicted value from predict()
```

### Display the depth of the trees
```{r}
xgb.plot.deepness(model = fit_coll$finalModel)
```

# Classification

## 1 tree
```{r}
set.seed(123)
ctrl = trainControl(method = "cv", number = 5)
private_simple = train(Private ~ ., 
             data = College,
             method = "xgbTree",
             tuneGrid = expand.grid(nrounds = 1, max_depth = 1:2, eta = 0.3, gamma = 0, colsample_bytree = .8, min_child_weight = 1, subsample = 1),
             trControl = ctrl)

private_simple
```

```{r}
xgb.plot.tree(model = private_simple$finalModel)
```

```{r}
 # newdata argument is optional here, but seems necessary on the homework.

predictions_priv = predict(private_simple, 
                           newdata = College, 
                           type = "prob")

head(predictions_priv)
College %>%
  select(Outstate, Accept, Enroll) %>%
  head()
```

Abilene Christian University:  Value = -0.156000003
```{r}

a= 0 + exp(-0.156000003) # 0 is the log odds from the baseline (.5)
1/(1+a)
```

## Classification with 2 trees